Why do we Eval?\\[2mm]\Large A large-scale study of Eval usage in R

### 2020-11-26 inst/paper/main.tex
background

\vspace{2mm}\noindent\emph{Is Past Prologue?} Previous work
investigated how \eval is used in web programming, specifically in web
sites that use JavaScript~\cite{pldi10a}. In 2010, 17 of the largest
web site used the feature. In 2011, 82\% of the 10,000 most accessed
sites used \eval~\cite{ecoop11}.  Yet, the strings passed to \eval,
and their behaviors when executed, are far from random; it was shown
that when one can observe several calls to \eval, the ``shape'' of
future calls can be predicted with 97\% accuracy~\cite{oopsla12b}.
Overall, practical usage suggested that most reflective calls were
relatively harmless. While this backs up the soundiness squad's
approach, does it generalize to other application domains than web
programming and to other languages?

@@

\subsection{Eval Semantics}

@@

\subsection{Previous Work}
Originally, \eval was limited to interpreted languages, but with the
advent of just-in-time compilation this restriction was gradually
lifted.

Richards et al.~\cite{ecoop11} provided the first large-scale study of
the runtime behavior of \eval in JavaScript. They analyzed a corpus of
the 10,000 most popular websites. They used an instrumented web
browser to gather execution traces and performed a dynamic analysis.
Their results show that \eval is pervasive with 82\% of the most
popular websites using it. The reasons for using \eval include the
desire to load code on demand (to reduce the time it takes to display
a page), the deserialization of JSON data (about 37\% of all calls)
and lightweight meta-programming to customize web pages.  While many
uses \eval were legitimate, many were unnecessary and could be
replaced with equivalent and safer code.  They reported on a
categorization of inputs to \eval that covers the vast majority of
input strings.  Restricting themselves to \eval in which all named
variables refer to the global scope, many patterns could be replaced
by more disciplined code~\cite{oopsla12b}.  The limitation of that
work are that there was no measure of code coverage, so the numbers
presented can be a lower bound on the possible behaviors.  Furthermore
in 2011, JavaScript usage was quite different from today. Their work
captured very few libraries and none of the more recent Node.js
ecosystem.  More details about dynamic analysis of JavaScript can be
found in~\cite{liang}.

Wang et al.~\cite{wang} analyzed usage of dynamic features in 18
Python programs include dynamic code generation to find if they affect
file change-proneness.  They reported that files with dynamic features
are significantly more likely to be the subject of changes, than other
files.  Chen et al. looked at the correlation between code changes and
dynamic features, including \eval, in 17 Python
programs~\cite{chen}. They did not observe many uses of \eval.  Callau
et al.~\cite{oscar} performed an empirical study of the usage of
dynamic features in the 1,000 largest Smalltalk projects in the
Squeaksource repository, accounting for more than 4 million lines of
code~\cite{oscar}. While \eval itself is not present, Smalltalk has a
rich reflective interface that gives programmer great
flexibility. Their work was based on a static analysis.

Morandat et al.~\cite{ecoop12} had a short section on the usage of
\eval in R. They found the \eval function to be widely used in R code
with 8500 call sites in CRAN and 2.2 million dynamic calls. The 15
most frequent call sites account for 88\% of eval. The \c{match.arg}
function is the highest used one with 54\% of all calls. In the other
call sites, they saw two uses cases. The most common is the evaluation
of the source code of a promise retrieved by \c{substitute} in a new
environment. This is done in the \c{with} function. The other use case
is the invocation of a function whose name or arguments are determined
dynamically. For this purpose, R provides \c{do.call} and thus using
\eval is overkill.

>>>>

\vspace{2mm}\noindent\emph{Is Past Prologue?} Previous work
investigated how \eval is used in web programming, specifically in
websites that use JavaScript~\cite{pldi10a}. In 2010, 17 of the
largest website used the feature. In 2011, 82\% of the 10,000 most
accessed sites used \eval~\cite{ecoop11}.  Yet, the strings passed to
\eval, and their behaviors when executed, are far from random; it was
shown that when one can observe several calls to \eval, the ``shape''
of future calls can be predicted with 97\% accuracy~\cite{oopsla12b}.
Overall, practical usage suggested that most reflective calls were
relatively harmless. While this backs up the soundiness squad's
approach, does it generalize to other application domains than web
programming and to other languages?

@@

\subsection{Eval Semantics}
Originally, \eval was limited to interpreted languages, but with the
advent of just-in-time compilation this restriction was gradually
lifted.

@@

\subsection{Previous Work}

Richards et al.~\cite{ecoop11} provided the first large-scale study of
the runtime behavior of \eval in JavaScript. They dynamically analyzed
a corpus of the 10,000 most popular websites with an instrumented web
browser to gather execution traces.  They show that \eval is pervasive
with 82\% of the most popular websites using it. The reasons for its
use include the desire to load code on demand, deserialization of JSON
dataq and lightweight meta-programming to customize web pages.  While
many uses were legitimate, just as many were unnecessary and could be
replaced with equivalent and safer code.  They categorized inputs to
\eval so as to cover the vast majority of input strings.  Restricting
themselves to \eval in which all named variables refer to the global
scope, many patterns could be replaced by more disciplined
code~\cite{oopsla12b}.  The work did not measure code coverage, so the
numbers presented are a lower bound on the possible behaviors.
Furthermore, JavaScript usage in 2011 is likely different from today,
e.g. Node.js was not covered by Richards.  More details about dynamic
analysis of JavaScript can be found in~\cite{liang}.

Wang et al.~\cite{wang} analyzed use of dynamic features in 18 Python
programs to find if they affect file change-proneness.  Files with
dynamic features are significantly more likely to be the subject of
changes, than other files.  Chen et al. looked at the correlation
between code changes and dynamic features, including \eval, in 17
Python programs~\cite{chen}. They did not observe many uses of \eval.
Callau et al.~\cite{oscar} performed an empirical study of the usage
of dynamic features in 1,000 Smalltalk projects. While \eval itself is
not present, Smalltalk has a rich reflective interface that gives
programmer great flexibility. Their work was based on a static
analysis.  %% TODO:: Say someting meaningfull about OSCAR

Morandat et al.~\cite{ecoop12} had a short section on the usage of
\eval in R. They found the it widely used in R code with 8500 call
sites in CRAN and 2.2 million dynamic calls. The 15 most frequent call
sites account for 88\% of those. The \c{match.arg} function is the
highest used one with 54\% of all calls. In the other call sites, they
saw two uses cases. The most common is the evaluation of the source
code of a promise retrieved by \c{substitute} in a new environment;
e.g. as done in the \c{with} function. The other use case is the
invocation of a function whose name or arguments are determined
dynamically. For this purpose, R provides \c{do.call} and thus \eval
is overkill.
